---
title: "Dad's Top 100"
output: html_document
date: "2022-08-30"
---

```{r setup, include=FALSE}
library(tidyverse);library(plotly); library(paletteer)
knitr::opts_chunk$set(warning = F, message = F, echo = F)
```

# Get Data

```{r get_games, eval = FALSE}
library(nbastatR)
Sys.setenv("VROOM_CONNECTION_SIZE" = 131072 * 2)

games <-
  game_logs(
    seasons = 1959:2022, league = "NBA", result_types = "player",
    season_types = "Regular Season", nest_data = F,
    assign_to_environment = TRUE, return_message = TRUE
  )

player_season <- bref_players_stats(seasons = 2000:2022, tables = c("advanced", "totals", "per_game"))

arrow::write_parquet(games,"data/games.parquet")
arrow::write_parquet(player_season,"data/player_season.parquet")
arrow::write_parquet(df_nba_player_dict,"data/df_nba_player_dict.parquet")

```

Criteria:

- Includes most recent performance (*i.e. best 'current' player, currently*)
- Must have played at least 41 games in 2021-2022 season

```{r clean_df}
# games <- arrow::read_parquet("data/games.parquet")
player_season <- arrow::read_parquet("data/player_season.parquet")
# df_nba_player_dict <- arrow::read_parquet("data/df_nba_player_dict.parquet")

df <-
  player_season %>%
  select(
    yearSeason, idPlayerNBA, namePlayer, agePlayer, 
    slugPosition, slugTeamBREF, where(is.numeric)
  ) %>%
  select(
    -countTeamsPlayerSeason,
    -countTeamsPlayerSeasonPerGame,
    -countTeamsPlayerSeasonTotals,-minutes,
    -ends_with("Totals"),-starts_with("url")
  ) %>%
  rename(
    season = yearSeason, player_id = idPlayerNBA, player_name = namePlayer,
    player_age = agePlayer, player_pos = slugPosition, player_team = slugTeamBREF
  ) %>%
  select(
    season,starts_with("player_"), countGames, countGamesStarted, minutesPerGame,
    starts_with("pct"),ends_with("PerGame"),starts_with("ratio")
  ) %>%
  # Remove "made" per game, since this is accounted for in % stats
  select(-ends_with("mPerGame")) %>%
  # Keep most recent season
  filter(season == max(season)) %>%
  filter(countGames >= 41)

rm(player_season)

```

# What does it mean to be the best?

Since we're focusing on current "best", we don't need to ask questions about how long someone has maintained their level of play (e.g. over a career, over a few years, etc.).  A rookie has just as much potential to be considered the "current best" as a LeBron.

Even so, there are are many ways we could define "best" of the current crop of NBA players, e.g.:

- Best at scoring?
- Most well-rounded?
- Biggest impact on team wins?

# Best at Scoring?

You can be good at scoring, but not try often.  Or shoot a lot and miss more often, taking opportunities away from others.  These players shoot *AND* score a lot...

```{r}
dt <-
  df %>%
  mutate(
    adj = ptsPerGame * pctTrueShooting,
    rank = row_number(desc(adj)),
    top_100 = if_else(rank <= 100,"diamond","circle")
  )

p <-
  dt %>%
  ggplot(aes(x = ptsPerGame, y = pctTrueShooting, color = adj)) +
  geom_point(aes(text = player_name, shape = top_100)) +
  scale_color_paletteer_c(`"pals::ocean.deep"`,direction = -1) +
  theme_minimal()

ggplotly(p)
```

```{r}
dt %>%
  select(player_name,rank,adj,ptsPerGame,pctTrueShooting) %>%
  arrange(rank) %>%
  DT::datatable(rownames = F)
```


# Biggest impact on team wins?

Compare advanced stats to take into account offensive and defensive contributions to wins (*i.e. Win Shares Per 48 Minutes*) and a normalized box score estimate of the defensive and offensive points per 100 possessions (*i.e. Box Plus/Minus*).


```{r}
dt <-
  df %>%
  mutate(
    adj = (ratioWSPer48 + ratioBPM) / 2,
    rank = row_number(desc(adj)),
    top_100 = if_else(rank <= 100,"diamond","circle")
  )

p <-
  dt %>%
  ggplot(aes(x = ratioWSPer48, y = ratioBPM, color = adj)) +
  geom_point(aes(text = player_name, shape = top_100)) +
  scale_color_paletteer_c(`"pals::ocean.solar"`,direction = -1) +
  theme_minimal()

ggplotly(p)
```

```{r}
dt %>%
  select(player_name,rank,adj,ratioWSPer48,ratioBPM) %>%
  arrange(rank) %>%
  DT::datatable(rownames = F)
```




# Stats Dictionary

- `season`: Season, such that e.g. 2000 corresponds to the 1999-2000 season.
- `player_id`: Unique NBA player ID.
- `player_pos`: Principal position played, e.g. small forward (SF), center (C), etc.
- `player_age`: Player's age on February 1 of the season
- `player_team`:
- `countGames`: Number of games played during season.
- `countGamesStarted`: Number of games started during season.
- `minutesPerGame`: Minutes Played Per Game.
- `pctTrueShooting`: A measure of shooting efficiency that takes into account 2-point field goals, 3-point field goals, and free throws.
- `pct3PRate`: 3-Point Attempt Rate; Percentage of FG Attempts from 3-Point Range
- `pctFTRate`: Free Throw Attempt Rate; Number of FT Attempts Per FG Attempt
- `pctORB`: Offensive Rebound Percentage; An estimate of the percentage of available offensive rebounds a player grabbed while they were on the floor.
- `pctDRB`: Defensive Rebound Percentage; An estimate of the percentage of available defensive rebounds a player grabbed while they were on the floor.
- `pctTRB` Total Rebound Percentage; An estimate of the percentage of available rebounds a player grabbed while they were on the floor.
- `pctAST`: Assist Percentage; An estimate of the percentage of teammate field goals a player assisted while they were on the floor.
- `pctSTL`: Steal Percentage; An estimate of the percentage of opponent possessions that end with a steal by the player while they were on the floor.
- `pctBLK`: Block Percentage; An estimate of the percentage of opponent two-point field goal attempts blocked by the player while they were on the floor.
- `pctTOV`: Turnover Percentage; An estimate of turnovers committed per 100 plays.
- `pctUSG`: Usage Percentage; An estimate of the percentage of team plays used by a player while they were on the floor.
- `pctFG`: Field Goal Percentage
- `pctFG3`: 3-Point Field Goal Percentage
- `pctFG2`: 2-Point Field Goal Percentage
- `pctEFG`: Effective Field Goal Percentage; This statistic adjusts for the fact that a 3-point field goal is worth one more point than a 2-point field goal.
- `pctFT`: Free Throw Percentage
- `fgaPerGame`: Average Field Goal Attempts Per Game
- `fg3aPerGame`: Average 3-Point Field Goal Attempts Per Game
- `fg2aPerGame`: Average 2-Point Field Goal Attempts Per Game
- `ftaPerGame`: Average Free Throw Attempts Per Game
- `orbPerGame`: Average Offensive Rebounds Per Game
- `drbPerGame`: Average Defensive Rebounds Per Game
- `trbPerGame`: Average Total Rebounds Per Game
- `astPerGame`: Average Assists Per Game
- `stlPerGame`: Average Steals Per Game
- `blkPerGame`: Average Blocks Per Game
- `tovPerGame`: Average Turnovers Per Game
- `pfPerGame`: Average Personal Fouls Per Game
- `ptsPerGame`: Average Points Per Game
- `ratioPER`: Player Efficiency Rating; A measure of per-minute production standardized such that the league average is 15.
- `ratioOWS`: Offensive Win Shares; An estimate of the number of wins contributed by a player due to offense.
- `ratioDWS`: Defensive Win Shares; An estimate of the number of wins contributed by a player due to defense.
- `ratioWS`: Win Shares; An estimate of the number of wins contributed by a player.
- `ratioWSPer48`: Win Shares Per 48 Minutes; An estimate of the number of wins contributed by a player per 48 minutes (league average is approximately .100)
- `ratioOBPM`: Offensive Box Plus/Minus; A box score estimate of the offensive points per 100 possessions a player contributed above a league-average player, translated to an average team.
- `ratioDBPM`: Defensive Box Plus/Minus; A box score estimate of the defensive points per 100 possessions a player contributed above a league-average player, translated to an average team.
- `ratioBPM`: Box Plus/Minus; A box score estimate of the points per 100 possessions a player contributed above a league-average player, translated to an average team.
- `ratioVORP`: Value over Replacement Player; A box score estimate of the points per 100 TEAM possessions that a player contributed above a replacement-level (-2.0) player, translated to an average team and prorated to an 82-game season.

